---
title: "WQD7004 Group Assignment"
#author: "Team 5"
#date: "`r Sys.Date()`"
output:
  html_document:
    css: styles.css    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Group Members (Team 5)
- Siti Nur Liyana Roslan (23067122)  
- Karenina Kamila (23117951)  
- Nur Ridwana Mohd Rafix (24054165)  
- Bong Hui Xin (22089462)  
- Salvin A/L Ravindran (17167138)  

# Dataset
## Title: Video Game Reviews and Ratings
## Year: 2024
## Purpose: A Randomly Generated Dataset to Help Practice Machine Learning Skills
## Source: <a href= "https://www.kaggle.com/datasets/jahnavipaliwal/video-game-reviews-and-ratings" class="custom-link">https://www.kaggle.com/datasets/jahnavipaliwal/video-game-reviews-and-ratings</a>

# Introduction

This report analyzes video game reviews and provides insights on various factors such as user ratings, platform distribution, genre distribution, and more. We will explore the data using visualizations, summary statistics, and generate a word cloud from user reviews.

# Loading Necessary Libraries

```{r}
# Load necessary libraries
#install.packages(c("tidyverse", "ggplot2", "dplyr", "corrplot", "vcd", "wordcloud", "RColorBrewer", "tm"))
#install.packages(c("rpart","rpart.plot", "caret", "e1071", "randomForest"))
library(rpart)
library(e1071)
library(dplyr)
library(caret)
library(rpart.plot)
library(randomForest)
library(tidyverse)
library(ggplot2)
library(corrplot)
library(vcd)
library(wordcloud)
library(RColorBrewer)
library(tm)
```

# Load the dataset
```{r}
data <- read.csv("video_game_reviews.csv")

# Display summary statistics of the dataset
summary(data)

# Get the dimension of the dataset
dim(data)

# Get the structure of the dataset
str(data)
glimpse(data)
```

# Data cleaning
```{r}

# View the column names
colnames(data)

# Dropping unnecessary columns; based on our objectives and correlation (numerical)
data1 <- data[c("Game.Title", "User.Rating", "Price", "Platform", "Release.Year", "Genre", "Game.Length..Hours.", "User.Review.Text")]

# Changing all chr columns to factor
data1$Genre <- factor(data1$Genre)
data1$Platform <- factor(data1$Platform)
data1$User.Review.Text <- factor(data1$User.Review.Text)

# Checking and handling for missing values
sum(is.na(data1))

# Replacing missing values (dummy)
```

# EDA
```{r}
# After cleaning the data set
summary(data1)
summary(data1$User.Rating)
sd(data1$User.Rating)
mean(data1$User.Rating)

# To find the highest rated game
high <- data1[order(-data$User.Rating), ][1, ]

# To find the lowest rated game
low <- data1[order(desc(-data$User.Rating)), ][1, ]

# Calculate quantiles of user ratings
quantiles <- quantile(data1$User.Rating, probs = c(0.25, 0.5, 0.75))
cat("25th Percentile:", quantiles[1], "\n")
cat("Median:", quantiles[2], "\n")
cat("75th Percentile:", quantiles[3], "\n")
```

```{r}

# 1. Univariate Analysis
# Bar plot of number of games by genres
ggplot(data1, aes(x = Genre)) + geom_bar(fill = "red", color = "black") + labs(title = "Number of Games by Genres", x = "Game Genres", y = "Number of Games" )

# Bar plot of game count by platform
ggplot(data1, aes(x = Platform)) + geom_bar(fill = "lightblue", color = "black") + labs(title = "Number of Games by Platform", x = "Platform", y = "Number of Games")

# Histogram of user rating for games released in year 2022
y22 <- data1%>%filter(Release.Year==2022)
ggplot(y22, aes(x = User.Rating)) + geom_histogram(binwidth = 1, fill = "orange", color = "black") + labs(title = "User Rating by Games Released in 2022", x = "Games Released", y = "User Rating")
```

```{r}
# 2. Bivariate Analysis
# Line graph of number of games released over years
count_release <- data1 %>% count(Release.Year)
ggplot(count_release, aes(x = Release.Year, y = n)) + geom_point() + geom_line() + labs(title = "Number of Games Released Over Years", x = "Years", y = "Number of Games")
# The graph peaked at 2020 with the most number of games released
# Might be due to global pandemic that increased the demand for home entertainment With lockdowns and social
# distancing measures in place. Gaming through digital platforms had resonated with audiences because they 
# provided a sense of normalcy and connection in a socially isolated world.

# Line graph of average rating by release year
ggplot(data1, aes(x = Release.Year, y = User.Rating)) + stat_summary(fun = "mean", geom = "line", color = "orange") + labs(title = "Average User Rating by Release Year", x = "Release Year", y = "Average Rating")
# Although 2020 has the highest number of game releases it also received the lowest average ratings 
# This can be attributed to several factors, including the increased player base and harsher standards in a
# crowded market. Mass production of games releases might dilute its quality which caused some games may
# have struggled to stand out, and the average quality might have dropped due to oversaturated market.

# Scatter plot for User Rating vs. Game Length (Hours) in year 2020
v <- data1%>%filter(Release.Year==2020)
ggplot(v, aes(x = User.Rating, y = Game.Length..Hours.)) + geom_point() + geom_smooth(method = lm) + labs(title = "Correlation between User Rating and Game Length (Hours) in year 2020")
```


```{r}
# 3. Multivariate Analysis
# Correlation matrix with numerical data
b <- data1[sapply(data1, is.numeric)]
b <- cor(b)
corrplot(b, method = "color", main = "Correlation Matrix with Numerical Data")
```


```{r}
# Pairs plot between numerical values
pairs(data1[,c(2,4,5,7)], main="Pairs Plot")
```


```{r}
# 4. Correlation Analysis for Categorical Data
chisq.test(data1$Genre, data1$Platform)
# Genre and Platform are highly correlated to each other (p=0.9282)

chisq.test(data1$Genre, data1$User.Review.Text)
chisq.test(data1$Platform, data1$User.Review.Text)
# Meanwhile, User Review Text data is only weakly correlated to both Genre and Platform data

quality<- data[c("User.Rating", "Graphics.Quality", "Soundtrack.Quality", "Story.Quality")]
chisq.test(quality$User.Rating, quality$Graphics.Quality)
chisq.test(quality$User.Rating, quality$Soundtrack.Quality)
chisq.test(quality$User.Rating, quality$Story.Quality)
# All of the game qualities are positively correlated with User Rating with Story Quality obtaining the 
# highest correlation compared to Graphics Quality and Soundtrack Quality
```


```{r}
# 5. WORD CLOUD
# Visualizing user review texts with word cloud
reviews <- data1$User.Review
# Clean the empty spaces between the texts
reviews <- gsub("\\s+", " ", reviews)
# Split the text into individual words
words <- unlist(strsplit(reviews, "\\s+"))
# Remove common stopwords
stopwords <- c(stopwords("en"), "but", "and", "many")
words <- words[!tolower(words) %in% stopwords]
# Create a frequency table of the words
freq <- table(words)
# Generate the word cloud
wordcloud(names(freq), freq, min.freq = 1, scale = c(4, 0.5), colors = brewer.pal(8, "PuOr"), random.order = F)
```

# Prepare Data For Modelling (Classification)
###create new column "popular" to classify the rating of the games basen on the 70th percentile threshold and split the data into 70% training and 30% testing
```{r}
# Check the structure of the dataset
str(data)

# Add Popularity column
data$Popularity <- ifelse(data$User.Rating >= 35.1, "Popular", "Unpopular")
data$Popularity <- as.factor(data$Popularity)
data$User.Rating <- as.numeric(data$User.Rating)

#spliting the data into training 70% and testing 30%
set.seed(123)
train_index <- createDataPartition(data$Popularity, p=0.7, list=FALSE)
train_data <- data[train_index, ]
test_data <- data[-train_index, ]

message(
  "Initial: ", nrow(data), " rows.\n",
  "Train: ", nrow(train_data), " rows  \n",
  "Test: ", nrow(test_data), " rows ."
)
```

# Objective 1 :  Classify Popular and Unpopular Games

## Classification Algorithm 1 : Decision Tree Model
```{r}
# Build the Decision Tree model
dt_model <- rpart(Popularity ~ Price + Game.Length..Hours.,
                  data = train_data,
                  method = "class")
# Predict on the test data
dt_predictions <- predict(dt_model, newdata = test_data, type = "class")
# Visualize the Decision Tree
rpart.plot(dt_model)
# Confusion matrix to evaluate the model
confusionMatrix(dt_predictions, test_data$Popularity)
```

## Classification Algorithm 2 : SVM Model
```{r}
# Train the SVM model
svm_model <- svm(Popularity ~ Price + Game.Length..Hours.,
                 data = train_data,
                 type = "C-classification",
                 kernel = "linear")
# Predict on the test data
svm_predictions <- predict(svm_model, newdata = test_data)
# Confusion matrix to evaluate the SVM model
confusionMatrix(svm_predictions, test_data$Popularity)
```

# Prepare Data For Modelling (Regression)
###prepare data by splitting it to 70% training and 30% testing and User.Rating as the target variable
```{r}
#regression
set.seed(123)
train_index <- createDataPartition(data$User.Rating, p=0.7, list=FALSE)
train_data <- data[train_index, ]
test_data <- data[-train_index, ]

message(
  "Initial: ", nrow(data), " rows.\n",
  "Train: ", nrow(train_data), " rows  \n",
  "Test: ", nrow(test_data), " rows ."
)
```

# Objective 2 :  Predict The Future Rating of Games

## Regression Algorithm 1 : Random Forest Model

```{r}
train_data$User.Rating <- as.numeric(train_data$User.Rating)
test_data$User.Rating <- as.numeric(test_data$User.Rating)
# Train the Random Forest model
rf_model <- randomForest(User.Rating ~ Price + Game.Length..Hours.,
                         data = train_data,
                         importance = TRUE,
                         ntree = 50)
# Make predictions on the test data
rf_predictions <- predict(rf_model, newdata = test_data)
# Calculate RMSE and R-Square
rmse <- sqrt(mean((rf_predictions - test_data$User.Rating)^2))
rsq <- cor(rf_predictions, test_data$User.Rating)^2
mae <- mean(abs(rf_predictions - test_data$User.Rating))

# Visualize prediction using plot
rf <- data.frame(y_test = test_data$User.Rating, y_pred = round(rf_predictions,1))
subset_rf<- rf[1:35, ]
par(mar = c(5, 4, 4, 2))
plot(subset_rf$y_test, type = "l", col = "black", lwd = 2, xlab = "Index", ylab = "Value")
lines(subset_rf$y_pred, col = "blue", lwd = 2)
legend("topright", legend = c("Actual", "Predicted"), col = c("black", "blue"), lwd = 2)
title(main="Actual vs Predicted for Random Forest Regressor Model")
```

## Regression Algorithm 2 : Linear Regression Model
```{r}
# Fit the linear regression model
lr_model <- lm(User.Rating ~ Price + Game.Length..Hours.,
            data = train_data)

# Make predictions on the test set
lr_predictions <- predict(lr_model, newdata = test_data)

# Calculate RMSE and R-Square
rmse_lr <- sqrt(mean((lr_predictions - test_data$User.Rating)^2))
rsq_lr <- cor(lr_predictions, test_data$User.Rating)^2
mae_lr <- mean(abs(lr_predictions - test_data$User.Rating))

# Visualize prediction using plot
lr <- data.frame(y_test = test_data$User.Rating, y_pred = round(lr_predictions,1))
subset_lr<- lr[1:35, ]
par(mar = c(5, 4, 4, 2))
plot(subset_lr$y_test, type = "l", col = "black", lwd = 2, xlab = "Index", ylab = "Value")
lines(subset_lr$y_pred, col = "blue", lwd = 2)
legend("topright", legend = c("Actual", "Predicted"), col = c("black", "blue"), lwd = 2)
title(main="Actual vs Predicted for Linear Regression Model")

```

##comparing Random forest and Linear Regression 
```{r}
#compare model performance
evalution_table<-data.frame(
  Model = c('Random Forest', 'Linear Regression'),
  R_squared = c(rsq,rsq_lr),
  RMSE = c(rmse,rmse_lr),
  MAE = c(mae,mae_lr)
)
print(evalution_table)

```

# Conclusion

###According to the correlation checking, price and game length hour have a real correlation with rating. This implies that the value of upcoming rating games can be predicted using these factors. The best model, as determined by the modeling part, is:

## First Objective: Classify Popular and Unpopular Games
### Best model : SVM Model
### Accuracy: 96.09%

## Second Objective: Predict The Future Rating of Games
### Best model : Linear Regression Algorithm
### Root Mean Squared Error (RMSE): 1.160752
### R-squared (R²): 0.9761487
### MAE : 1.007670
