---
title: "Game Review"
author: "Team 5"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

This report analyzes video game reviews and provides insights on various factors such as user ratings, platform distribution, genre distribution, and more. We will explore the data using visualizations, summary statistics, and generate a word cloud from user reviews.

## Loading Necessary Libraries
```{r}
# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(corrplot)
library(vcd)
library(wordcloud)
library(RColorBrewer)
library(tm)
```


## Load the dataset
```{r}
data <- read.csv("video_game_reviews.csv")

# Display summary statistics of the dataset
summary(data)

# Get the dimension of the dataset
dim(data)

# Get the structure of the dataset
str(data)
glimpse(data)
```

## Data cleaning
```{r}

# View the column names
colnames(data)

# Dropping unnecessary columns; based on our objectives and correlation (numerical)
data1 <- data[c("Game Title", "User Rating", "Platform", "Price", "Release Year", "Genre", "Game Length (Hours)", "User Review")]

# Changing all chr columns to factor
data1$Genre <- factor(data1$Genre)
data1$Platform <- factor(data1$Platform)
data1$`User Review` <- factor(data1$`User Review`)

# Checking and handling for missing values
sum(is.na(data1))

# Rename column names


```

## EDA
```{r}
# After cleaning the data set
summary(data1)
summary(data1$User_Rating)
sd(data1$User_Rating)
mean(data1$User_Rating)

# To find the highest rated game
high <- data1[order(-data$User_Rating), ][1, ]

# To find the lowest rated game
low <- data1[order(desc(-data$User_Rating)), ][1, ]


# 1. Univariate Analysis
# Bar plot of number of games by genres
ggplot(data1, aes(x = Genre)) + geom_bar(fill = "red", color = "black") + labs(title = "Number of Games by Genres", x = "Game Genres", y = "Number of Games" )

# Bar plot of game count by platform
ggplot(data1, aes(x = `Platform`)) + geom_bar(fill = "lightblue", color = "black") + labs(title = "Number of Games by Platform", x = "Platform", y = "Number of Games")

# Histogram of user rating for games released in year 2022
y22 <- data1%>%filter(`Release Year`==2022)
ggplot(y22, aes(x = `User Rating`)) + geom_histogram(binwidth = 1, fill = "orange", color = "black") + labs(title = "User Rating by Games Released in 2022", x = "Games Released", y = "User Rating")


# 2. Bivariate Analysis
# Line graph of number of games released over years
count_release <- data1 %>% count(`Release Year`)
ggplot(count_release, aes(x = `Release Year`, y = n)) + geom_point() + geom_line() + labs(title = "Number of Games Released Over Years", x = "Years", y = "Number of Games")

# Line graph of average rating by release year
ggplot(data1, aes(x = `Release Year`, y = `User Rating`)) + stat_summary(fun = "mean", geom = "line", color = "orange") + labs(title = "Average User Rating by Release Year", x = "Release Year", y = "Average Rating")

# Scatter plot for User Rating vs. Game Length (Hours) in year 2020
v <- data1%>%filter(`Release Year`==2020)
ggplot(v, aes(x = `User Rating`, y = `Game Length (Hours)`)) + geom_point() + geom_smooth(method = lm) + labs(title = "Correlation between User Rating and Game Length (Hours) in year 2020")


# 3. Multivariate Analysis
# Correlation matrix with numerical data
b <- cor(b)
corrplot(b, method = "color", main = "Correlation Matrix with Numerical Data")

# Pairs plot between numerical values
pairs(data1[,c(2,4,5,7)], main="Pairs Plot")


# 4. WORD CLOUD
# Visualizing user review texts with word cloud
reviews <- data1$`User Reviews`
# Clean the empty spaces between the texts
reviews <- gsub("\\s+", " ", reviews)
# Split the text into individual words
reviews <- strsplit(reviews, "\\s+")
# Remove common stopwords
reviews <- c(stopwords("en"), "but", "and", "many", "game")
reviews <- words[!words %in% reviews]
reviews <- table(reviews)
wordcloud(names(reviews), reviews, min.freq = 1, scale = c(4, 0.3), colors = brewer.pal(8, "PuOr"), random.order = F)
```

## Modelling
```{r}
# 1. Classification

# 2. Regression
```

## Model Evaluation
```{r}
```

## Conclusion
```{r}
